Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2025-01-13 manuscript/manuscript.tex
minor edits to section 2
Minor edits. Grammar.

\begin{abstract}
In the early 21st century, the open data movement began to transform 
societies and governments by promoting transparency,
innovation, and public engagement. The City of New York (NYC) has been at
the forefront of this movement since the enactment of the Open 
Data Law in 2012, leading to the creation of the NYC Open Data
portal. This portal currently hosts 2,700 datasets,
serving as a crucial resource for research across various domains, 
including health, urban development, and transportation. However, the
effective use of open data relies heavily on data quality and
usability, challenges that remain insufficiently addressed in the
literature. This paper examines these challenges through a
case study of the NYC 311 Service Request dataset,  identifying key
issues in data validity, consistency, and curation efficiency. We
propose a set of data curation principles, tailored for
government-released open data, to address these challenges.
Our findings highlight the importance of harmonized field definitions,
streamlined storage, and automated quality checks, offering practical
guidelines for improving the reliability and utility of open datasets.
\end{abstract}

@@

Our investigation uses a 311 SR dataset covering  2-years (2022-2023), 
and was collected on 15 September 2024. It is to conduct 
analysis in areas such as data consistency, data validity, 
redundant fields, etc.. Instructions on downloading this file are 
available in the supplemental material. Characteristics of the 311 SR dataset include:

@@

\item A free-form text field, \texttt{resolution\_description}, which 
  supports 930 characters including commas and 
  special characters; often problematic for automated processing.
\end{itemize}
\medskip

@@

The Mostly Empty category includes:
\texttt{taxi\_company\_borough}, 
\texttt{bridge\_highway\_name},
\texttt{due\_date}, and \texttt{road\_ramp}. 
The Partially Empty category includes:
\texttt{cross\_street\_1 \& \_2},
\texttt{location\_type}, and texttt{landmark}. And 
the Few/None Empty category includes \texttt{created\_date}, 
\texttt{complaint\_type},
\texttt{agency}, 
and \texttt{status}. 
Of particular note is that several of the Mostly Empty fields are used by only one or 
two Agencies (See: \autoref{tab:field-usage-summary}).

@@

\begin{table}[tbp]
\centering
\caption{Sample field usage by select agencies among the 6.4 million
  SRs in 2022--2023. }
\label{tab:field-usage-by-agency}
\small
\begin{tabular}{lrrrrrrr}
\toprule
\textbf{\textbf{Field}} & \textbf{DEP} & \textbf{DHS} & \textbf{DOB} & \textbf{DOT} & \textbf{DPR} & \textbf{DSNY} & \textbf{TLC} \\
\midrule
\texttt{landmark}                  & 2935     & 70185    & 282      & 135150   & 254218   & 424284   & 63133  \\
\texttt{facility\_type}            & 175549   & 0        & 51575    & 159667   & 0        & 66012    & 0      \\
\texttt{due\_date}                 & 0        & 0        & 0        & 0        & 0        & 32097    & 0      \\
\texttt{community\_board}          & 361715   & 99613    & 203715   & 401143   & 266201   & 608188   & 73601  \\
\texttt{vehicle\_type}             & 0        & 0        & 0        & 0        & 0        & 0        & 1696   \\
\texttt{taxi\_company\_borough}    & 0        & 0        & 0        & 0        & 0        & 0        & 4025   \\
\texttt{taxi\_pick\_up\_location}  & 0        & 0        & 0        & 0        & 0        & 0        & 72558  \\
\texttt{bridge\_highway\_name}     & 15       & 21012    & 0        & 7451     & 0        & 613      & 977    \\
\texttt{bridge\_highway\_direction} & 15      & 8973     & 0        & 7450     & 0        & 614      & 1006   \\
\texttt{road\_ramp}                & 15       & 2588     & 0        & 7450     & 0        & 613      & 999    \\
\texttt{bridge\_highway\_segment}  & 15       & 21013    & 0        & 7450     & 0        & 613      & 987    \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{abstract}
In the early 21st century, the open data movement began to transform 
societies and governments by promoting transparency,
innovation, and public engagement. The City of New York (NYC) has been at
the forefront of this movement since the enactment of the Open 
Data Law in 2012, creating the NYC Open Data
portal. The portal currently hosts 2,700 datasets,
serving as a crucial resource for research across various domains, 
including health, urban development, and transportation. However, the
effective use of open data relies heavily on data quality and
usability, challenges that remain insufficiently addressed in the
literature. This paper examines these challenges via a
case study of the NYC 311 Service Request dataset,  identifying key
issues in data validity, consistency, and curation efficiency. We
propose a set of data curation principles, tailored for
government-released open data, to address these challenges.
Our findings highlight the importance of harmonized field definitions,
streamlined storage, and automated quality checks, offering practical
guidelines for improving the reliability and utility of open datasets.
\end{abstract}

@@

Our investigation uses a 311 SR dataset covering  2-years (2022-2023), 
and was collected on 15 September 2024. It is to conduct 
analysis in areas such as data consistency, data validity, 
redundant fields, etc.. Instructions on downloading this file are 
available in the supplemental material. Characteristics of this
two-year 311 SR dataset are as follows.

@@

\item A free-form text field, \texttt{resolution\_description}, which 
  supports 930 characters including commas and 
  special characters; often problematic for automated processing.
\end{itemize}

@@

The Mostly Empty category includes:
\texttt{taxi\_company\_borough}, 
\texttt{due\_date}, 
\texttt{road\_ramp},
and \texttt{bridge\_highway\_name}. 
The Partially Empty category includes:
\texttt{location\_type}, 
\texttt{landmark} and \texttt{cross\_street\_1 \& \_2}.
The Few/None Empty category includes \texttt{created\_date}, 
\texttt{complaint\_type},
\texttt{agency}, 
and \texttt{status}. 
Note that several of the Mostly Empty fields are used by only one or 
two Agencies (See: \autoref{tab:field-usage-summary}).

@@

\begin{table}[tbp]
\centering
\caption{Sample field usage by select agencies among the 6.4 million
  SRs in 2022--2023. }
\label{tab:field-usage-by-agency}
\small
\begin{tabular}{lrrrrrrr}
\toprule
\textbf{\textbf{Field}} & \textbf{DEP} & \textbf{DHS} & \textbf{DOB} & \textbf{DOT} & \textbf{DPR} & \textbf{DSNY} & \textbf{TLC} \\
\midrule
\texttt{landmark}                  & 2935     & 70185    & 282      & 135150   & 254218   & 424284   & 63133  \\
\texttt{facility\_type}            & 175549   & 0        & 51575    & 159667   & 0        & 66012    & 0      \\
\texttt{due\_date}                 & 0        & 0        & 0        & 0        & 0        & 32097    & 0      \\
\texttt{vehicle\_type}             & 0        & 0        & 0        & 0        & 0        & 0        & 1696   \\
\texttt{taxi\_company\_borough}    & 0        & 0        & 0        & 0        & 0        & 0        & 4025   \\
\texttt{taxi\_pick\_up\_location}  & 0        & 0        & 0        & 0        & 0        & 0        & 72558  \\
\texttt{bridge\_highway\_name}     & 15       & 21012    & 0        & 7451     & 0        & 613      & 977    \\
\texttt{road\_ramp}                & 15       & 2588     & 0        & 7450     & 0        & 613      & 999    \\
\bottomrule
\end{tabular}
\end{table}
